Categorical and gradient aspects of wordlikeness\thanks{Thanks to Gene Buckley, Charles Yang, and audiences at NELS 43 for helpful comments.}

### 2012-11-29 4.tex
Brought 4 up to speed with CLS 47 paper changes.

\noindent
It is not necessary for predictable prosodic structure to be present in underlying form to achieve this effect, contrary to what has sometimes been stated \citep[e.g.,][255]{A74}.
All that is necesary is that no phonological processes intervene between underlying representation and the surface form of medial clusters.
Under the further assumption that underlying and surface representations are as similar as alternations allow (e.g., \citealt[205f.]{Dell1973}, \citealt[28f.]{Stampe1973}), this is the null hypothesis.

@@

\begin{table}
\centering
\begin{tabular}{l rrrr}
\toprule
           & attested & unattested & saturation & $p$-value \\
\midrule
conforming & 25       & 91         & 22\%       & \multirow{2}{*}{.106} \\
violating  &  4       & 40         &  9\%       \\
\bottomrule
\end{tabular}
\caption{Dorsal-labial cluster attestation in the lexical sample}
\label{dltab}
\end{table}

@@

\begin{table}
\centering
\begin{tabular}{l rrrr}
\toprule
           & attested & unattested & saturation & $p$-value \\
\midrule
conforming & 47       & 512        &  8\%       & \multirow{2}{*}{.250} \\
violating  &  0       &  25        &  0\%       \\
\bottomrule
\end{tabular}
\caption{ABA cluster attestation in the lexical sample}
\label{abatab}
\end{table}

@@

\begin{table}
\centering
\begin{tabular}{l rrrr}
\toprule
           & attested & unattested & saturation & $p$-value \\
\midrule
conforming & 35       & 329        & 10\%       & \multirow{2}{*}{.002} \\
violating  & 11       & 305        &  3\%       \\
\bottomrule
\end{tabular}
\caption{Obstruent voice assimilation cluster attestation in the lexical sample}
\label{ovatab}
\end{table}

@@

\begin{table}
\centering
\begin{tabular}{l rrrr}
\toprule
           & attested & unattested & saturation & $p$-value \\
\midrule
conforming & 31       & 2          & 94\%       & \multirow{2}{*}{3.4\e{-07}} \\
violating  & 11       & 22         & 33\%       \\
\bottomrule
\end{tabular}
\caption{Nasal place assimilation cluster attestation in the lexical sample}
\label{npatab}
\end{table}

@@

\noindent
As can be seen from Table \ref{degemtab}, the sample includes no sequences of identical segments, or of identical segments differing only in voice, something highly unlikely to arise by chance.

\begin{table}
\centering
\begin{tabular}{l rrrr}
\toprule
           & attested & unattested & saturation & $p$-value \\
\midrule
conforming & 173      & 643        & 21\%       & \multirow{2}{*}{1.2\e{-10}} \\
violating  & 0        & 104        & 0\%        \\
\bottomrule
\end{tabular}
\caption{Degemination in the lexical sample}
\label{degemtab}
\end{table}

@@

\begin{table}
\centering
\begin{tabular}{l | rrrr}
\toprule
                    & accuracy & precision & recall & $F_1$ \\
\midrule
Baseline            & 0.812    & 0.812     & 1.000  & 0.896 \\
Expected frequency  & 0.836    & 0.845     & 0.984  & 0.909 \\
Derived constraints & 0.838    & 0.840     & 0.997  & 0.912 \\
DC \& EF            & 0.861    & 0.868     & 0.988  & 0.924 \\
\citet{Hayes2008a}  & 0.835    & 0.942     & 0.854  & 0.896 \\
\bottomrule
\end{tabular}
\caption{Results for the cluster classification task; a combination of derived constraints and expected frequency (``DC \& EF'') produces the highest accuracy and $F_1$.}
\label{cmresults}
\end{table}

@@

\paragraph{Maximum entropy phonotactics} 
\citet{Hayes2008a} present a model which uses the principle of maximum entropy to weigh a large number of competing phonotactic constraints. 
In one sense, this is isomorphic to the expected frequency/derived constraint model in that the constraint discovery mechanism is sensitive to the expected frequency of clusters: it favors constraints which rule out clusters with high expected frequency (but which are unattested) over those which have low expected frequency.
Unlike the expected frequency/derived constraint model, however, the constraints discovered are not limited to those which are illustrated by alternations; it may also posit static constraints, much like those proposed by \citet{Pierrehumbert1994}.
Since this model has many experimenter-determined parameters, 
an attempt is made to replicate the details of \citeauthor{Hayes2008a}'s evaluation as closely as is feasible: their software, model parameters, and feature specifications are all used. 
Following the procedure of \citet{HayesInPress}, a novel feature [$\pm$\textsc{Coda}] is added to allow the model to distinguish coda and onset consonants. 
Since the maximum entropy model produces slightly different scores on each run, the worst-performing of 10 runs is reported here, following \citet{Hayes2008a}.
This model has the poorest recall of any model; compared to the derived constraints baseline, the constraints induced by the maximum entropy model are narrower.
This is particularly clear regarding possible clusters with a nasal coda followed by a non-homorganic obstruent, like *[m.kl]: the vast majority of such clusters, which would be neutralized by \textsc{Nasal Place Assimilation}, are unattested, but many are erroneously assigned the highest possible score by the maximum entropy model.

@@

\begin{figure}
\centering
\includegraphics{co.pdf}
\caption{Medial coda and medial onset type frequencies in the lexical sample show a Zipfian distribution; frequencies have been smoothed using the $Z_r$ transform (see Appendix \ref{zr}).}
\label{cao}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{sim.pdf}
\caption{A simulated cluster inventory, generated by random sampling and applying phonological neutralizations closely matches the observed cluster frequencies (represented by the smoothing line). Simulated and observed frequencies have been smoothed using the $Z_r$ transform (see Appendix \ref{zr}).}
\label{sim}
\end{figure}

>>>>

\noindent
It is not necessary for predictable prosodic structure to be present in underlying form to achieve this effect, contrary to what has sometimes been stated \citep[e.g.,][255]{A74}.
All that is necesary is that no phonological processes intervene between underlying representation and the surface form of medial clusters.
This is little more than the definition of a ``faithful'' cluster.
Under the further assumption that underlying and surface representations are as similar as phonological processes allow (e.g., \citealt[205f.]{Dell1973}, \citealt[28f.]{Stampe1973}), faithful realization is the null hypothesis.

@@

\begin{table}[ht]
\centering
\begin{tabular}{l rrrr}
\toprule
           & attested & unattested & saturation & $p$-value \\
\midrule
conforming & 25       & 91         & 22\%       & \multirow{2}{*}{.106} \\
violating  &  4       & 40         &  9\%       \\
\bottomrule
\end{tabular}
\caption{Dorsal-labial cluster attestation in the lexical sample}
\label{dltab}
\end{table}

@@

\begin{table}[ht]
\centering
\begin{tabular}{l rrrr}
\toprule
           & attested & unattested & saturation & $p$-value \\
\midrule
conforming & 47       & 512        &  8\%       & \multirow{2}{*}{.250} \\
violating  &  0       &  25        &  0\%       \\
\bottomrule
\end{tabular}
\caption{ABA cluster attestation in the lexical sample}
\label{abatab}
\end{table}

@@

\begin{table}[ht]
\centering
\begin{tabular}{l rrrr}
\toprule
           & attested & unattested & saturation & $p$-value \\
\midrule
conforming & 35       & 329        & 10\%       & \multirow{2}{*}{.002} \\
violating  & 11       & 305        &  3\%       \\
\bottomrule
\end{tabular}
\caption{Obstruent voice assimilation cluster attestation in the lexical sample}
\label{ovatab}
\end{table}

@@

\begin{table}[ht]
\centering
\begin{tabular}{l rrrr}
\toprule
           & attested & unattested & saturation & $p$-value \\
\midrule
conforming & 31       & 2          & 94\%       & \multirow{2}{*}{3.4\e{-07}} \\
violating  & 11       & 22         & 33\%       \\
\bottomrule
\end{tabular}
\caption{Nasal place assimilation cluster attestation in the lexical sample}
\label{npatab}
\end{table}

@@

\noindent
As can be seen from Table \ref{degemtab}, the sample includes no sequences of identical segments, or of identical segments differing only in voice, something highly unlikely to arise by chance.
Stampean occultation by \textsc{Degemination} provides a natural explanation for this gap.
It is informative to compare this to the constraint against ``ABA'' clusters proposed by \citeauthor{Pierrehumbert1994}: both are exceptionless, but only the former is unlikely to arise by chance.

\begin{table}[ht]
\centering
\begin{tabular}{l rrrr}
\toprule
           & attested & unattested & saturation & $p$-value \\
\midrule
conforming & 173      & 643        & 21\%       & \multirow{2}{*}{1.2\e{-10}} \\
violating  & 0        & 104        & 0\%        \\
\bottomrule
\end{tabular}
\caption{Degemination in the lexical sample}
\label{degemtab}
\end{table}

@@

\begin{table}[ht]
\centering
\begin{tabular}{l | rrrr}
\toprule
                    & accuracy & precision & recall & $F_1$ \\
\midrule
Baseline            & 0.812    & 0.812     & 1.000  & 0.896 \\
Expected frequency  & 0.836    & 0.845     & 0.984  & 0.909 \\
Derived constraints & 0.838    & 0.840     & 0.997  & 0.912 \\
DC \& EF            & 0.861    & 0.868     & 0.988  & 0.924 \\
\citet{Hayes2008a}  & 0.835    & 0.942     & 0.854  & 0.896 \\
\bottomrule
\end{tabular}
\caption{Results for the cluster classification task; a combination of derived constraints and expected frequency (``DC \& EF'') produces the highest accuracy and $F_1$.}
\label{cmresults}
\end{table}

@@

\paragraph{Maximum entropy phonotactics} 
\citet{Hayes2008a} present a model which uses the principle of maximum entropy to weigh a large number of competing phonotactic constraints. 
In one sense, this is isomorphic to the expected frequency/derived constraint model in that the constraint discovery mechanism is sensitive to the expected frequency of clusters: it favors constraints which rule out clusters with high expected frequency (but which are unattested) over those which have low expected frequency.
Unlike the expected frequency/derived constraint model, however, the constraints discovered are not limited to those which are illustrated by alternations; it may also posit static constraints, much like those proposed by \citet{Pierrehumbert1994}.

Since this model has numerous experimenter-defined parameters, a close replication of \citeauthor{Hayes2008a}'s original paper is attempted: both their implementation and phonological feature specifications are used here.
Following \citet{HayesInPress}, dictionary entries are syllabified using the procedure described in \S\ref{syllabification} and a novel feature [$\pm$
\textsc{Coda}] is added to allow the model to distinguish coda and onset co
nsonants.
Also, following \citeauthor{Hayes2008a}, constraints are limited to those s
panning as many as three segments and an ``accuracy schedule'' of [$.001$, 
$.01$, $.1$, $.2$, $.3$] is used.
Since the maximum entropy model produces slightly different scores on each 
run, the worst-performing of 10 runs is reported here, following \citeauthor{Hayes2008a}.
This model has the poorest recall of any model; compared to the derived constraints baseline, the constraints induced by the maximum entropy model are narrower.
This is particularly clear regarding possible clusters with a nasal coda followed by a non-homorganic obstruent, like *[m.kl]: the vast majority of such clusters, which would be neutralized by \textsc{Nasal Place Assimilation}, are unattested, but many are erroneously assigned the highest possible score by the maximum entropy model.

@@

\begin{figure}[ht]
\centering
\includegraphics{co.pdf}
\caption{Medial coda and medial onset type frequencies in the lexical sample show a Zipfian distribution; frequencies have been smoothed using the $Z_r$ transform (see Appendix \ref{zr}).}
\label{cao}
\end{figure}

@@

\begin{figure}[ht]
\centering
\includegraphics{sim.pdf}
\caption{A simulated cluster inventory, generated by random sampling and applying phonological neutralizations closely matches the observed cluster frequencies (represented by the smoothing line). Simulated and observed frequencies have been smoothed using the $Z_r$ transform (see Appendix \ref{zr}).}
\label{sim}
\end{figure}
